### 2025-03-12 paper/preprint/paper.tex
Add final subsection to discussion

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{plain}
\newtheorem{corollary}{Research Question}[section]
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\textbf{\emph{Interventions on features may have downstream impacts on
fairness.}}\\
- \href{https://arxiv.org/abs/2401.16088}{Equality of opportunity}?\\
- \href{https://proceedings.mlr.press/v202/gao23d/gao23d.pdf}{Social
segregation}?\\
- Supporting AR within one context/system may still unfairly target
recourse-affected individuals within other contexts/systems? In Example
3.1 younger individuals gain access to a loan but their age could still
lead to, e.g., enforcing stronger supervisory mechanisms?

@@

@@

@@

@@

>>>>

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{corollary}{Research Question}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\textbf{\emph{Interventions on features may have downstream impacts on
fairness.}} Related to the point above, we provide a tool that allows
practitioners to modify the sensitivity of a model with respect to
certain features, which may have implication for the fair and equitable
treatment of individuals subject to automated decisions. As protecting a
set of features leads the model to assign higher relative importance to
unprotected features, model owners could misuse our solution by
enforcing explanations based on features that are more difficult to
modify by some (group of) individuals. For example, consider again the
Adult dataset where features such as workclass or education may be more
difficult to change for underpriviledged groups. When applied
irresponsibly, counterfactual training could result in an unfairly
assigned burden of recourse (e.g.,
\citeproc{ref-sharma2020certifai}{Sharma, Henderson, and Ghosh 2020}),
threatening the equality of opportunity in the system
(\citeproc{ref-bell2024fairness}{Bell et al. 2024}) and potentially
reinforcing social segregation (\citeproc{ref-gao2023impact}{Gao and
Lakkaraju 2023}). Still, as the referenced publications indicate, such
phenomena are not specific to CT; regrettably, all types of ADM
solutions without strong external protections have been recognized to
promote harmful power dynamics (\citeproc{ref-maas2023machine}{Maas
2023}).

@@

\bibitem[\citeproctext]{ref-bell2024fairness}
Bell, Andrew, Joao Fonseca, Carlo Abrate, Francesco Bonchi, and Julia
Stoyanovich. 2024. {``Fairness in Algorithmic Recourse Through the Lens
of Substantive Equality of Opportunity.''}
\url{https://arxiv.org/abs/2401.16088}.

@@

\bibitem[\citeproctext]{ref-gao2023impact}
Gao, Ruijiang, and Himabindu Lakkaraju. 2023. {``On the Impact of
Algorithmic Recourse on Social Segregation.''} In \emph{Proceedings of
the 40th International Conference on Machine Learning}. ICML'23.
Honolulu, Hawaii, USA: JMLR.org.

@@

\bibitem[\citeproctext]{ref-maas2023machine}
Maas, Jonne. 2023. {``Machine Learning and Power Relations.''} \emph{AI
\& SOCIETY} 38 (4): 1493--1500.

@@

\bibitem[\citeproctext]{ref-sharma2020certifai}
Sharma, Shubham, Jette Henderson, and Joydeep Ghosh. 2020. {``CERTIFAI:
A Common Framework to Provide Explanations and Analyse the Fairness and
Robustness of Black-Box Models.''} In \emph{Proceedings of the AAAI/ACM
Conference on AI, Ethics, and Society}, 166--72. AIES '20. New York, NY,
USA: Association for Computing Machinery.
\url{https://doi.org/10.1145/3375627.3375812}.

### 2025-03-12 paper/sections/discussion.qmd
Add final subsection to discussion

***Interventions on features may have downstream impacts on fairness.***  
- [Equality of opportunity](https://arxiv.org/abs/2401.16088)?  
- [Social segregation](https://proceedings.mlr.press/v202/gao23d/gao23d.pdf)?  
- Supporting AR within one context/system may still unfairly target recourse-affected individuals within other contexts/systems? In Example 3.1 younger individuals gain access to a loan but their age could still lead to, e.g., enforcing stronger supervisory mechanisms?

>>>>

***Interventions on features may have downstream impacts on fairness.***  Related to the point above, we provide a tool that allows practitioners to modify the sensitivity of a model with respect to certain features, which may have implication for the fair and equitable treatment of individuals subject to automated decisions. As protecting a set of features leads the model to assign higher relative importance to unprotected features, model owners could misuse our solution by enforcing explanations based on features that are more difficult to modify by some (group of) individuals. For example, consider again the Adult dataset where features such as workclass or education may be more difficult to change for underpriviledged groups. When applied irresponsibly, counterfactual training could result in an unfairly assigned burden of recourse [e.g., @sharma2020certifai], threatening the equality of opportunity in the system [@bell2024fairness] and potentially reinforcing social segregation [@gao2023impact]. Still, as the referenced publications indicate, such phenomena are not specific to CT; regrettably, all types of ADM solutions without strong external protections have been recognized to promote harmful power dynamics [@maas2023machine].
